---
title: A Regression Analysis on The Efects of The Risk of Infection and on The Length
  of Stay in Hospitals
author:
- Rumil Legaspi, Rumil.legaspi@gmail.com
- Mei Leng Lao, Email
date: "28 February 2021"
output:
  word_document: default
  toc: yes
    toc_depth: 2
  html_notebook:
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 2
    latex_engine: xelatex
  theme: lumen
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
---

---------------------------------------------------------------------
# __Purpose__

We are conducting a simple linear regression model from the SENIC dataset to analyze the relationship of the explanatory variable, infection of risk(INFRISK) and the response variable, length of stay(LOS).


# __Our Data__

## Quick background on Dataset and variable

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
#Setting up our work environment
setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Project 1/Project 1")

#Loading in packages
library(nortest)
library(car)
library(lmtest)
library(formatR)

#Loading in the data
load(file = "SENIC.rdata")

Infection_data <- data.frame("SENIC.rdata")

#Defining and renaming our Explanatory(X) and Response(Y) variables
infection_risk = SENIC$INFRISK #X
length_of_stay = SENIC$LOS #Y
age = SENIC$AGE #Z
```
<p style="text-align: center;">
__(delete this)__
some interpretations:

- Length of stay is explained by the average estimated probability of acquiring infection in hospital. 

- As the risk of infection increases the average length of stay in the hospital also increases.
</p>

---------------------------------------------------------------------

# __Part 1: Interpretation and Parameter Inference__

## Estimated Linear Regression Function

```{r, tidy=TRUE}
#Generating our Linear Model using lm() then summarizing
infection_lm = lm(length_of_stay ~ infection_risk, data = Infection_data)
summary(infection_lm)
```


<div align="center">
From summarizing our Linear Regression model we can see:

__$\beta_{0}$ = 6.3368 _(intercept)_ __

__$\beta_{1}$ = 0.7604 _(slope)_ __

and the estimated regression equation to be:

__$\hat{Y}= 6.3368 + 0.7604X$__
</div>

## Fitting on Scatterplots

```{r fig.align="center"}
plot(SENIC$INFRISK, SENIC$LOS)
abline(infection_lm, col = "red")
```
## Interpretting Regression Coefficients & $R^2$

__From our model we derive that our intercept, $\beta_{0}$ = 6.3368__:

This indicates where our response output lies when there is no input or when $X$ is 0. In other words, when risk of infection (explanatory variable) is at 0, the average length of stay of patients in a hospital is roughly 6 days. 

Analyzing the intercept on its own might be confusing and at times misleading. In understanding the context of our data we can see that despite patients having an average estimated probability of acquiring an infection in a hospital be 0% we know that this is impossible. Additionally we know that it is possible for patients to be in the hospital for roughly 6 days for other medical reasons.

In other words, although a bit misleading at first glance, when risk of infection is close to zero and almost nonexistent, there is still truth in a patient having a prolonged length of stay in a hospital.


__Our slope, $\hat{\beta_1} = 0.7604$__:

Indicates as the risk of infection increases by 1 unit, the average length of stay increases by 0.74 days. This can also be thought of as when the risk of infection increases by 1% the average length of stay in a hospital increases by about 18 hours.

__Our $R^2=0.2846$__:

***The R squared found at $0.2846$ indicates that the risk of infection (input variable) helps explain close to 28% of the variability in our response variable, length of stay. In other words, __the linear relationship between our model and the dependent variable length of stay is fairly weak.__


## Hypothesis Testing on our Slope to Test Significance

Our null hypothesis is that there is no linear relationship 

<div align="center">

__Null Hypothesis__: ${H_0}$: $\beta_{1} = 0$ (slope is horizontal/ no relationship), in other words there is no linear relationship between risk of infection and length of stay

__Alternative Hypothesis__: ${H_1}$: $\beta_{1} \neq 0$ (slope exists/ relationship exists), there is linear relationship either positive or negative between risk of infection and length of stay.
      
</div>

### Testing Using the p-value

The slope indicates a positive relationship and the p-value (1.177e-09) is very close to 0 which is less than our $\alpha =0.05$, this indicates that we can reject the null hypothesis and conclude with the alternative hypothesis ***that the linear relationship is significant.

### Conducting a T-test ***(is this even necessary?)

From our summary we have the t-value of the risk of infection(explanatory variable) $t= 6.645$

***Now to compare this with a critical value we must first find it
 Using T-table to find critical value:
 - N = 113, So our degrees of freedom is 111
 - one tail test with $\alpha$ = 0.05 and with 95% prob


```{r}
#Using qt() in R we get

qt(0.975, 111)
```
Our critical value is 1.981567

Now comparing the absolute value of the t-statistic of $t= 6.645$ to the slope ($\beta_{1}$) of risk of infection with our critical value 1.981567 

<div align="center">
      
  |6.645| > 1.981567    
      
</div>

we can conclude that because the absolute value of our t-statistic is greater than our critical value, we reject the null hypothesis and conclude with our alternative hypothesis that __a relationship exists between risk of infection(x) and length of stay(Y).__


### Finding the 95% Confidence Interval of the Slope

```{r}
#alpha at 0.05
alpha <- 0.05

#constructing our 95% confidence interval
confint(infection_lm, level = 1 - alpha)
```
#### Interpretation


This output reads that within our confidence interval from 2.5% (the lower limit of our interval) to 97.5% (the upper limit of our interval), our intercept and __slope__ are both found within the listed intervals.

In this case if we repeat this experiment many times, the true population parameter of our slope $\beta_{1}$  will be between the interval 0.5336442 and 0.9871976 with 95% confidence and $\alpha$ (accepted error) of 5%

0 is not included in our interval, but we are interested in it because if zero was included in our confidence interval then that would indicate (that there is a chance that) no change/relationship and would make risk of infection(INRFRISK) a bad predictor for length of stay(LOS). So in this case, since 0 is not included, we can conclude that there is change. 

---------------------------------------------------------------------

# __Part 2: Point and interval estimation__

## Conducting 95% Confidence Interval when Length of Stay (Input Variable) is 5 for the Mean Length of Stay (Response Variable)

```{r}
#Creating a new single observation where risk is 5
new_infection_data <- data.frame(infection_risk = 5)

#Constructing our prediction interval
ci_infection_5 <- predict(infection_lm, new_infection_data,
                          interval = "confidence", level = 1 - alpha )

ci_infection_5
```
## Confidence Interval Interpretation when INFRISK = 5

The fitted value of the length of stay variable when the risk of infection __is at 5% is 10.13889.__

This 95% confidence interval when risk of infection __is at 5% is from 9.802 to 10.475.__ 

In other words, when the risk of infection is 5%, we can expect the __average(true mean)__ of the length of stay (response variable) to be within the intervals of __roughly 9 to 10 and a half days with 95% confidence.__

## Constructing a Prediction Interval

We can use a prediction interval when trying to find where an individual observation will fall. Lets construct a prediction interval given risk of infection is at 5 percent.

```{r}
#Constructing prediction interval when INFRISK is 5
pi_infection_5 <- predict(infection_lm, new_infection_data,  interval = "prediction", level = 1 - alpha )

pi_infection_5
```
## Prediction Interval Interpretation

***From the results we can predict with 95% confidence that when a patient has a risk of infection at 5%, her length of stay will fall somewhere between 6.903 and 13.37 days. (should I say roughly 7 days to 13 days?)

---------------------------------------------------------------------

# __Part 3: Diagnostics__

## Our Assumptions

To ensure our model is still within the bounds of our made assumptions for a linear regression model lets plot them using different plotting methods.

lets recall our made assumptions for a linear regression model:

- __L__ inearity
- __I__ ndependence
- __N__ ormality of the errors
- __E__ qual error variance for all values of X (homoskedasticity)



```{r}
#plotting scatterplots to check assumption
plot(infection_lm , which = c(1))
```
### ***Testing Linearity

We cannot see a clear violation of linearity assumption in our __residual vs fitted plot__ since we do not see a systematic pattern.  



### ***Independence
residuals vs time

```{r}
#Checking to independence

#install.packages("MASS")
#library(MASS)
#infection.resid = studres(infection_lm)

#Studentized residuals vs. predictor
#plot(SENIC$infection_risk, infection.resid)

#Residuals vs. Order
#data.order = c(1:25)
#plot(data.order, )
```


## Testing Normality

### Plotting to Check for Normality and Equal Variance Assumptions

```{r}
#Plotting a boxplot and #histogram
boxplot(infection_lm$residuals)
hist(infection_lm$residuals)
```

## Boxplot and Histogram Interpretation
We can see that our boxplot is fairly symmetrical and our histogram shows our residuals somewhat resemble a bell shape.
Thus, so far our normality assumption still holds true.

### Stating our Hypothesis

<div align="center">

__Null Hypothesis__: ${H_0}$: The data __is__ from a normal distribution

__Alternative Hypothesis__: ${H_1}$: The data is __NOT__ from a normal distribution
      
</div>

### Testing our Hypothesis

__To test these we can use several normality tests using...__

- Shapiro-Wilk normality test
- Shapiro-Francia normality test
- Anderson-Darling normality test

These tests focus mainly on the usage of regression residuals with an p-value as an output useful for hypothesis testing. __Are main goal is to see if our data truly follows a normal distribution.__

```{r}
#Shapiro-Wilk normality test
shapiro.test(infection_lm$residuals)

#Shapiro-Francia normality test
nortest::sf.test(infection_lm$residuals)

#Anderson-Darling normality test
nortest::ad.test(infection_lm$residuals)
```

## Interpretation of Normality Tests

Looking at the results of these three tests we can see that the p-values are small. Therefore we cannot reject our NULL hypothesis and that __there is an issue with our normality assumption.__

### Plotting to Check for Equal Variance Assumptions

## Testing Equal Variance Assumptions



### Stating our Hypothesis

<div align="center">

__Null Hypothesis__: ${H_0}$:  The variances in the data __is__ equal
__Alternative Hypothesis__: ${H_1}$: The variances in the data are __NOT__ equal
      
</div>

### ***Testing our Hypothesis

We dont want to see fan shapes otherwise they violate equal variance assumption

we want equal spread
```{r}
#Residuals vs predictor variable
plot(infection_lm$residuals, SENIC$infection_risk)

#Conducting Levene Test
#bf.test(infection_lm, data = SENIC)

```
Residuals vs predictor variable plot indicates equal spread therefore equal variances of error (homoskedasticity) is not violated.


## Checking for omitted predictors

```{r}
#scatterplot
plot(SENIC$AGE, SENIC$LOS)
plot(SENIC$INFRISK, SENIC$LOS)

#Plotting infection_lm model residuals vs Age
plot(SENIC$AGE, infection_lm$residuals)
```

In the residuals vs. the potentially omitted variable (Age) the plots are randomly scattered and show no particular kind of relation between the residuals and Age.
```{r}
age_lm <- lm(age ~ infection_risk, data = Infection_data)
summary(age_lm)
```
```{r}
qt(0.975, 111)

```
Because the absolute value of our critical value is less than our T-value we fial to reject our NULL hypothesis.


Age is not a potentially omitted variable.

---------------------------------------------------------------------
