---
title: 'STAT 511: Week 12 Lecture '
author: "Rumil Legaspi"
date: "5 April 2021"
output:
  pdf_document: 
    latex_engine: xelatex
  html_notebook: default
  html_document:
    df_print: paged
  word_document: default
---

# Multiple Regression & Brand Preference Dataset

### Setting up workspace

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
library(nortest)
library(olsrr)
library(car)
library(lmtest)
library(MASS)
library(tidyverse)

setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Week 12")

#reading in a csv
#brand_data = read.table(file = "Brand.txt", header = FALSE, sep = "")

#reading in a r data file
load(file = "bodyfat.Rdata")

bodyfat <- data.frame(data)


#View(brand_data)

# #Adding headers
#names(brand_data) <- c("Rating", "Moisture", "Sweetness")

# names(bank_data) <- c("", "")

#Defining dependent and independent vars
#Rating = _data$ #Y
#Moisture = _data$ #X1
#Sweetness = _data$ #X2
```
 
 
 to what extend is the sum of sqaure errors reduction or increase (of r^2?). We want to test on that amount. (watch video)
 
Setting up regression models before plugging into anova function for f test
```{r}


#Defining dependent and independent vars
fat = bodyfat$bodyfat #Y
triceps = bodyfat$triceps #X1
thigh = bodyfat$thigh #X2
arm = bodyfat$arm

triceps_lm <- lm(fat ~ triceps, data = bodyfat)
thigh_lm <- lm(fat ~ thigh, data = bodyfat)
triceps_thigh_lm <- lm(fat ~ triceps + thigh, data = bodyfat)
arm_lm <- lm(fat ~ arm, data = bodyfat)
full_lm <- lm(fat ~ triceps + thigh + arm, data = bodyfat)

```
Doing Partial F test
stating null and alternative

null: there is no change when adding certain predictors to the significance of our model
h
alt: there is change when adding certain predictors towards the significance of our model


```{r}
#Comparing tricep only with tricep + thigh (thigh missing)
anova(triceps_lm, triceps_thigh_lm)
```
tricep is significant and we reject null hypothesis indicating that when thigh is missing and we are left with tricep only our model remains significant with a alpha of 0.05. aka thigh is a significant variable

whichever variable is extra that means its that variable that is being tested

```{r}
#Comparing thigh only with tricep + thigh (tricep missing)
anova(thigh_lm, triceps_thigh_lm)
```
We fail to reject null hpothesis since p value is greater than our alpha indicating when tricep is missing our model is not significant with an accepted error(alpha) pf 0.05. suggesting thigh is not significant predictor variable

```{r}
#comparing
anova(arm_lm, full_lm)
```
since this is very significant means tricep and thigh are jointly significant


# Scaled Regression
```{r}
#Scaling coefficients
scaled_fat <- scale(fat)
scaled_triceps <- scale(triceps)
scaled_thigh<-scale(thigh)
scaled_arm<-scale(arm)
```
```{r}
#putting scaled coeffs into lm now the results are scaled, w/o intercept
scaled_lm <- lm(scaled_fat ~ 0 + scaled_triceps + scaled_thigh + scaled_arm, data = bodyfat)
scaled_lm
summary(scaled_lm)
```
Therefore when holding outher variables constant/unchanged when _____ variable increases by ____ standard deviations we can expect _____ to increase by _____ standard deviations.


▪ Regress fat on all three predictors
```{r}
fat_lm <- lm(fat ~ triceps + thigh + arm, data = bodyfat)
fat_lm
summary(fat_lm)
```
notice how above there are no significant predictors but overall model test is highly significant




▪ Regress scaled fat on scaled three predictors
```{r}
scaled_fat_lm <- lm(scaled_fat ~  scaled_triceps + scaled_thigh + scaled_arm, data = bodyfat)
fat_lm
summary(fat_lm)
```

 Regress scaled fat on scaled three predictors without the intercept
To exclude intercept in a regression model, use lm(y ~ 0 + x1+ x2).

```{r}
scaled_lm <- lm(scaled_fat ~ 0 + scaled_triceps + scaled_thigh + scaled_arm, data = bodyfat)
scaled_lm
summary(scaled_lm)
```

 
 
 