---
title: "STAT 511 Week 10 Lecture"
author: "Rumil Legaspi"
date: "3/22/2021"
output: pdf_document
---



```{r, collapse = TRUE, warning = FALSE, message = FALSE}
library(nortest)
library(olsrr)
library(car)
library(lmtest)
library(MASS)
library(tidyverse)

setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Week 10")

dstudio_data = read.table(file = "dstudios.txt", header = FALSE, sep = "")

View(dstudio_data)

# #Adding headers
names(dstudio_data) <- c("targtpop", "dispoinc", "sales")

# names(bank_data) <- c("", "")

#Defining dependent and independent vars
sales = dstudio_data$sales #Y
dispoinc = dstudio_data$dispoinc #X1
targtpop = dstudio_data$targtpop #X2
```

```{r}
sales.lm = lm(sales ~ targtpop + dispoinc)
summary(sales.lm)

```
From summarizing our  Multiple Linear Regression model we can see:

and the estimated regression equation to be:

__$\hat{Y}= -68.8571 + -1.4546X + 9.3655X$__


__$\beta_{0}$ = -68.8571 _(intercept)_ __

interpretation: When target population and disposable income is both 0 the sales of the portait will be bankrupt and be at a deficit of -68 thousand dollars. Makes sense and does not makes sense

__$\beta_{1}$ = 1.4546 _(slope)_ __

Interpretation: holding dispoinc constand and unchanged 

__$\beta_{2}$ = 9.3655 _(slope)_ __

Interpretation: Holding target popo constant and unchanged when disploable income increase by 1 thousand dollars sales increase by 1 thousand dollars

residuals variance equal residual std err but squared

#R Squared and Adjusted R Squared, Constructing a Confidence interval
```{r}
summary(sales.lm)
```
1. Multiple R^2 = 0.9167; Adjusted R^2 = 0.9075
 
Interpretation: Multiple R^2 explains to us that 91% of the variability in Y is collectively explained by our two X variables, disposable income and target population.
 
Adjust R^2 tells us that 90& of the variability in Y is collectively explained by our two variables while penalizes explanatory variables that are redundant and irrelevant.

2. Conducting Test for overall model significance

```{r}
#alpha at 0.05
alpha <- 0.05


#constructing our 95% confidence interval
confint(sales.lm, level = 1 - alpha)
confint(sales.lm)
```

__Null Hypothesis__: ${H_0}$: $\beta_{1}=\beta_{2} = 0$ (slopes are showing no change), in other words show no overall model significance between our partial slopes.

__Alternative Hypothesis__: ${H_1}$: $\beta_{1}=\beta_{2} \neq 0$ (slopes are showing change), in other words show overall model significance between our partial slopes.

Using P value or critical value (F Statistic larger than critical value)
P-value = 1.921e-10
F Statistic = 99.1






