---
title: 'STAT 511: Analyzing Home Sales Prices Using Multiple Regression Analysis'
author: 
- Rumil Legaspi, Rumil.legaspi@gmail.com
- Efe Umukoro, Eumukoro20@apu.edu
- Solange Ebobisse Mapenya, bebobissemapenya20@apu.edu
date: "4/28/2021"
output: pdf_document
toc: yes
toc_depth: 2
geometry: margin=1in
header-includes:
 \usepackage{float}
---

```{r setup, include=FALSE, echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE,tidy.opts=list(width.cutoff=60), tidy = TRUE)
```

# **Background & Objective**

Given that a city tax assessor is interested in predicting residential home sales prices in a midwestern city with various characteristics, we will be conducting a **multiple linear regression analysis (MLR)** from the Real Estate Sales (APPENC07) dataset from 2002. We aim to observe and predict the relationships using the given features, ***square feet***, the absence or presence of a ***swimming pool*** and ***air conditioning***, and our response variable as ***house sales price***.

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
#Setting up our work environment
library(tidyverse)
library(formatR)

#Loading in the text data
raw_data = read.table(file = "C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Project 2/APPENC07.txt", header = FALSE, sep = "")

#Converting into tibble data frame for easier data analysis
house_data <- as_tibble(raw_data)
```

```{r}
#Defining and renaming our Explanatory(X) and Response(Y) variables
house_data <- house_data %>% select(sales_price = V2,
                                    square_feet = V3,
                                    swimming_pool = V8,
                                    air_conditioning = V6)

#Setting explanatory and response variables
sales_price <-  house_data %>% select(sales_price) #Y
square_feet <- house_data %>% select(square_feet) #X1
swimming_pool <- house_data %>% select(swimming_pool) #X2
air_conditioning <- house_data %>% select(air_conditioning) #X3
```

### Our dataset is comprised of *522 total transactions* from midwestern home sales during the year 2002.

```{r}
#Glimpse of our dataset
head(house_data)
```

# Part 1 - Regression using a Dummy Variable

## 1a. Estimated regression equation from regressing sales price on swimming pool only.

```{r}
#Regressing sales price only on swimming pool dummy variable
pool_lm <- lm(sales_price ~ swimming_pool, data = house_data)

#summarizing linear model
summary(pool_lm)
```

### Estimated Regression model:

$\hat{Y}= 272396 + 79724X$

## 1b. Interpretation of estimated intercept and slope.

### Intercept: $B_{0}$ = 272396

The estimated mean Y-value when X = 0 (reference/baseline group) is 272396. When put in context, the mean sales price of a house when the property **does not** contain a swimming pool is estimated to be \$272,396.

### Slope: $B_{1}$ = 79724

The slope of 79724 in our model indicates the change for the sales price of a property **containing** a swimming pool, **relative** to a property **without** a swimming pool to be \$352,120.

The calculations of these coefficients can be represented in this table.

+--------------------------------+------------------------------+-----------------------------+
| $\hat{Y}= B_{0} + B_{1}X_{1}$  | Swimming Pool = No           | Swimming Pool = Yes         |
+================================+==============================+=============================+
| $\hat{Y}= 77.375 + 8.750X$     | $\hat{Y}= 272396 + 79724(0)$ | $\hat{Y}= 272396+ 79724(1)$ |
+--------------------------------+------------------------------+-----------------------------+
|                                | = 272396                     | = 272396+ 79724             |
+--------------------------------+------------------------------+-----------------------------+
| **Estimated Mean Sales Price** | **\$272,396**                | **\$352,120**               |
+--------------------------------+------------------------------+-----------------------------+

: Property Sales Price With & Without Swimming Pool

## 1c. Hypothesis test on the significance of the slope coefficient.

Using a significance level of $\alpha=0.05$.

**Null Hypothesis:** ${H_0}$: $\beta_{j} = 0$ (slopes are showing no change), $X_{j}$ is not linearly associated with Y, therefore the partial slope is not significant.

**Alternative Hypothesis**: ${H_1}$: $\beta_{j} \neq 0$ (slopes are showing change), $X_{j}$ is linearly associated with Y, therefore the partial slope is significant.

Testing the significance of a property **with** a swimming pool ($\hat{\beta_1} = 79724$)

Conclusion and Decision Rule using p-value:

Because the **p-value** for having a swimming pool is [1] 0.00078 and is significantly smaller than $\alpha=0.05$, we reject our NULL hypothesis and conclude that our partial slope, that a property **containing** a swimming pool in reference to one **without a swimming pool**, shows statistical significance in our model.

\newpage

# Part 2 - Fitting a MLR model With the Interaction Term of a Dummy and Continuous Variable

## 2a. Regressing sales price on the (1) swimming pool dummy variable, (2) area of residence, and the (3) interaction between these two variables.

```{r}
pool_sqft_lm <- lm(sales_price ~ swimming_pool + 
                                square_feet +
                              swimming_pool * square_feet,
                                      data = house_data)
summary(pool_sqft_lm)
```

Estimated regression equation for each kind of property:

$\hat{Y}= -88538.996 + 105909.972X + 161.910Y - 37.213(X*Y)$

*Variable Assignment:*

**X** = Swimming pool

**Y** = Square feet

**X \* Y** = Interaction of swimming pool and square feet

+-----------------------------------------------------------+-------------------------------------------------------------+-------------------------------------------------------------+
| $\hat{Y}= B_{0} + B_{1}X_{}+B_{2}Y+B_{3}(X*Y)$            | Swimming Pool = No                                          | Swimming Pool = Yes                                         |
+===========================================================+=============================================================+=============================================================+
| $\hat{Y}= -88538.996 + 105909.972X+161.910Y- 37.213(X*Y)$ | $\hat{Y}= -88538.996 + 105909.972(0)+161.910Y- 37.213(0*Y)$ | $\hat{Y}= -88538.996 + 105909.972(1)+161.910Y- 37.213(1*Y)$ |
+-----------------------------------------------------------+-------------------------------------------------------------+-------------------------------------------------------------+
|                                                           |                                                             |                                                             |
+-----------------------------------------------------------+-------------------------------------------------------------+-------------------------------------------------------------+
|                                                           | $= -88538.996 +161.910Y$                                    | $= -88538.996 + 105909.972+161.910Y- 37.213(Y)$             |
+-----------------------------------------------------------+-------------------------------------------------------------+-------------------------------------------------------------+
|                                                           |                                                             | $= 17370.976 + 124.697Y$                                    |
+-----------------------------------------------------------+-------------------------------------------------------------+-------------------------------------------------------------+
|                                                           |                                                             |                                                             |
+-----------------------------------------------------------+-------------------------------------------------------------+-------------------------------------------------------------+
| **Estimated Regression Equations**                        | $= -88538.996 +161.910Y$                                    | $= 17370.976 + 124.697Y$                                    |
+-----------------------------------------------------------+-------------------------------------------------------------+-------------------------------------------------------------+

: Calculating Estimated Regression Equations for Properties With and Without Pools

\newpage

## 2b. Plotting fitted regression lines

```{r}
# Code to plot regression equations model:
plot_coef <- house_data %>% 
  
  ggplot(aes(x = square_feet, y = sales_price, color = as.factor(swimming_pool))) +
 
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Square Feet", 
       y = "Sales Price", 
       color = "Swimming Pool",
       title = "Fitted Regression Lines Plotted",
       subtitle = "Sales Prices of Properties With & Without Pools By Square Feet") +
  scale_y_continuous(labels = scales::dollar)
  
plot_coef

```

To find the value at which these two lines intersect algebraically we can set both equations equal to one another, solving for one variable, then plugging that back into the equation to get the other variable to obtain the coordinates.

The point of intersection of these two lines are when the values of: **Square feet is 2846.04 and sales price = 372264.58**.

## 2c. Testing if the two regression lines are parallel.

Using a significance level of $\alpha=0.05$.

Null Hypothesis: ${H_0}$: $\beta_{c} = 0$ Partial slope of the interaction is 0.

Alternative Hypothesis: ${H_1}$: $\beta_{c} \neq 0$ Partial slope of the interaction is not 0.

Testing the significance of a property **with** a swimming pool ($\hat{\beta_3} = -37.213$)

Conclusion and Decision Rule using p-value:

Looking at our model summary, we see that the p-value of our interaction term is [1]0.0300 which means we **reject** NULL hypothesis and \*\*\*conclude with our alternative hypothesis and that our regression lines are not parallel and a relationship exists between the two lines (because the interaction coefficient is not 0).

# Part 3 - MLR Only With the Interaction of Dummy Variables

## 3a. Fitting a MLR on both swimming pool and AC dummy variables and find the estimated regression equation.

```{r}
#Regressing house sales price on both dummy variables: swimming pool and AC, plus their interaction


pool_ac_lm <- lm(sales_price ~ swimming_pool + 
                              (air_conditioning) +
                               swimming_pool * air_conditioning,
                               data = house_data)

summary(pool_ac_lm)
```

### Estimated regression equation for each kind of property:

$\hat{Y}= 189578.2 + 421.8X + 100875.8Z - 65876.5(X*Z)$

*Variable Assignment:*

**X** = Swimming pool

**Z** = Air conditioning

**X \* Z** = Interaction of swimming pool and air conditioning

## 3c. Calculating estimated mean sales prices for 4 types of properties using estimated regression equation:

1.  No swimming pool and no AC

```{r}
#Coding in our variables based on its absence or presence
x = 0
z = 0

#Running our regression formula with our coded variables
reg_eq_1 <- 189578.2 + 421.8*x + 100875.8*z + 65876.5*x*z

#printing
reg_eq_1

```

The estimated mean sales price of a property without a swimming pool and AC is \$189,578.2.

2.  No swimming pool and has AC

```{r}
#Coding in our variables based on its absence or presence
x = 0
z = 1

#Running our regression formula with our coded variables
reg_eq_2 <- 189578.2 + 421.8*x + 100875.8*z + 65876.5*x*z

#printing
reg_eq_2
```

The estimated mean sales price of a property without a swimming pool but has AC is \$290,454.

3.  Has swimming pool and no AC

```{r}
#Coding in our variables based on its absence or presence
x = 1
z = 0

#Running our regression formula with our coded variables
reg_eq_3 <- 189578.2 + 421.8*x + 100875.8*z + 65876.5*x*z

#printing
reg_eq_3
```

The estimated mean sales price of a property with a swimming pool and no AC is \$190,000.

4.  Has swimming pool and has AC

```{r}
#Coding in our variables based on its absence or presence
x = 1
z = 1

#Running our regression formula with our coded variables
reg_eq_4 <- 189578.2 + 421.8*x + 100875.8*z + 65876.5*x*z

#printing
reg_eq_4
```

The estimated mean sales price of a property with a swimming pool and AC is \$356,752.3.

# Conclusion

Based on our regression analysis it is clear that a property with a swimming pool and air conditioning (\$356,752.3), cost significantly more than a property without (\$189,578.2.).

For future reference we understand that our dataset is unbalanced with only about 7%, or 36 out of 522 observations owning swimming pools and 16%, or 88 out of 522 observations having air conditioning. Moving forward one way to correct this would be to collect more data from houses owning these features.
