---
title: 'STAT 592: Project 3'
author: 
- Rumil Legaspi, Rumil.legaspi@gmail.com
- Efe Umukoro, email
- Solange Ebobisse Mapenya, email
date: "4/28/2021"
output: pdf_document
toc: yes
toc_depth: 2
geometry: margin=1in
header-includes:
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Background & Objective**

Given that a city tax assessor is interested in predicting residential home sales prices in a midwestern city with various characteristics, we will be conducting a **multiple linear regression analysis** from the Real Estate Sales (APPENC07) dataset from 2002. We aim to observe and predict the relationship using the given features, ***square feet***, the absence or presence of a ***swimming pool*** and ***air conditioning***, and our response variable as ***house sales price***.

```{r, collapse = TRUE, warning = FALSE, message = FALSE}

#Setting up our work environment
setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Project 2")
library(nortest)
library(olsrr)
library(car)
library(lmtest)
library(MASS)
library(tidyverse)
library(ggcorrplot)
#Loading in the text data
raw_data = read.table(file = "APPENC07.txt", header = FALSE, sep = "")

#Converting into tibble data frame for easier data analysis
house_data <- as_tibble(raw_data)
```

```{r}
#Defining and renaming our Explanatory(X) and Response(Y) variables
house_data <- house_data %>% select(sales_price = V2,
                                    square_feet = V3,
                                    swimming_pool = V8,
                                    air_conditioning = V6)

#Setting explanatory and response variables
sales_price <-  house_data %>% select(sales_price) #Y
square_feet <- house_data %>% select(square_feet) #X1
swimming_pool <- house_data %>% select(swimming_pool) #X2
air_conditioning <- house_data %>% select(air_conditioning) #X3

```

# Part 1 - Regression using a Dummy Variable

## 1a. Estimated regression equation from regressing sales price on swimming pool only

```{r}
#Regressing sales price only on swimming pool dummy variable
pool_only <- lm(sales_price ~ swimming_pool, data = house_data)

#summarizing linear model
summary(pool_only)
```

## 1b. Interpretation of estimated intercept and slope.

### Intercept: $B_{0}$ = 272396

The estimated mean Y-value when X = 0 (reference/baseline group) is 272396. When put in context, the mean sales price of a house when the property **does not** contain a swimming pool is estimated to be \$272,396.

### Slope: $B_{1}$ = 79724

The slope of 79724 in our model indicates the change for the sales price of a property **containing** a swimming pool, **relative** to a property **without** a swimming pool to be \$352,120.

The calculations of these coefficients can be represented in this table.

+-------------------------------+------------------------------+-----------------------------+
| $\hat{Y}= B_{0} + B_{1}X_{1}$ | Swimming Pool = No           | Swimming Pool = Yes         |
+===============================+==============================+=============================+
| $\hat{Y}= 77.375 + 8.750X$    | $\hat{Y}= 272396 + 79724(0)$ | $\hat{Y}= 272396+ 79724(1)$ |
+-------------------------------+------------------------------+-----------------------------+
|                               | = 272396                     | = 272396+ 79724             |
+-------------------------------+------------------------------+-----------------------------+
| Estimated Mean Sales Price    | **\$272,396**                | **\$352,120**               |
+-------------------------------+------------------------------+-----------------------------+

: Property Sales Price With & Without Swimming Pool

# Part 2 - MLR with the Interaction Term of a Dummy and Continuous Variable

# Part 3 - MLR Only with the Interaction of Dummy Variables
