---
title: 'STAT 511: HW #7'
author: "Rumil Legaspi"
date: "4/19/2021"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    latex_engine: xelatex
---

# Multiple Regression & Brand Preference Dataset

### Setting up workspace

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
library(nortest)
library(olsrr)
library(car)
library(lmtest)
library(MASS)
library(tidyverse)

setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Week 10/Week 10/")

brand_data = read.table(file = "Brand.txt", header = FALSE, sep = "")

View(brand_data)

# #Adding headers
names(brand_data) <- c("Rating", "Moisture", "Sweetness")

# names(bank_data) <- c("", "")

#Defining dependent and independent vars
Rating = brand_data$Rating #Y
Moisture = brand_data$Moisture #X1
Sweetness = brand_data$Sweetness #X2
```

# 1a. Regress degree of brand liking on sweetness only. Write down the estimated regression model.

```{r}
#Coding Sweetness as a dummy variable
cat_sweetness <- as.factor(Sweetness)

#Regressing Rating on new Sweetness dummy variable
Sweetness_only <- lm(Rating ~ cat_sweetness, data = brand_data)

summary(Sweetness_only)
```

Estimated Regression model:

$\hat{Y}= 77.375 + 8.750X$

```{r}
#Checking how many levels there are in sweetness as a dummy
levels(cat_sweetness)
typeof(cat_sweetness)
attributes(cat_sweetness)
```

So in this case since sweetness is a category with only 2 level (sweetness level 2 and sweetness level 4) we can think of our regression model like so:

We will use sweetness level 2 as our reference group

```{r}
#the "2" indicates the level marked as "2" given from our level() function and now setting sweetness level 2 as our reference group
cat_sweetness_new <- relevel(cat_sweetness, ref = "2")

#Regressing
Sweetness_only2 <- lm(Rating ~ cat_sweetness_new, data = brand_data)

summary(Sweetness_only2)
```

## 1b. Compute the estimated mean degree of brand liking at each level of sweetness, i.e., what is

the estimated mean degree of brand liking at sweetness level 2? At sweetness level 4?

After reading the summary we can now say the estimated means when Sweetness level is 2 and 4 to be.

+-------------------------------+------------------------------+------------------------------+
| $\hat{Y}= B_{0} + B_{1}X_{1}$ | Sweetness = Level 2          | Sweetness = Level 4          |
+===============================+==============================+==============================+
| $\hat{Y}= 77.375 + 8.750X$    | $\hat{Y}= 77.375 + 8.750(0)$ | $\hat{Y}= 77.375 + 8.750(1)$ |
+-------------------------------+------------------------------+------------------------------+
|                               | = 77.375                     | = 77.375 + 8.750             |
+-------------------------------+------------------------------+------------------------------+
|                               | 77.375                       | 86.125                       |
+-------------------------------+------------------------------+------------------------------+

: Mean Rating at Level 2 or Level 4 Sweetness

# 1c. Interpret the intercept coefficient.

$B_{0}$ = 77.375

The estimated mean Y-value when X = 0 which is our reference/baseline group. When put in context, this is the mean Rating when the brand's sweetness level is 2.

# 1d. Interpret the slope coefficient.

The change in mean Rating for Sweetness level 4 relative to sweetness level 2 is 86.125 (77.375 + 8.750(1)).

## 1e. Is the slope coefficient significant? State the null, alternative, decision rule and conclusion.

**Null Hypothesis**: ${H_0}$: $\beta_{j} = 0$ (slopes are showing no change), $X_{j}$ **is not** linearly associated with Y, therefore the partial slope is not significant.

**Alternative Hypothesis**: ${H_1}$: $\beta_{j} \neq 0$ (slopes are showing change), $X_{j}$ **is** linearly associated with Y, therefore the partial slope is significant.

#### Testing the significance of Sweetness Level 4 ($\hat{\beta_1} = 8.750$)

**p-value:**

Because the p-value for sweetness level 4 is [1] 0.13 and is greater than our alpha (accepted error/significance level) of 0.05 we **fail to reject** our NULL hypothesis and conclude that our partial slope, **Sweetness Level 4** in reference to Sweetness Level of 2, does not show significance (\*\*\*in our model).

```{r}
#***t.test(Sweetness, cat_sweetness_new)
```

# 2. Refer to the "Brand Preference" dataset. Code sweetness ($X_{2}$) as a dummy variable.

## 2a. Fit a multiple regression model with moisture content, sweetness, and their interaction.

```{r}
#Regressing Rating on Moisture and with Sweetness still as a dummy variable
full_lm <- lm(Rating ~ cat_sweetness + Moisture, data = brand_data)

summary(full_lm)
```

## 2b. Write down the estimated regression equation at each sweetness level

$\hat{Y}= 46.40 + 8.75X + 4.425X$

Sweetness level 2: $\hat{Y}= 46.40 + 4.425X$

Sweetness level 4: $\hat{Y}= 46.40 + 8.75X + 4.425X$

## \*\*\*2c. Interpret the slope coefficient in each estimated regression equation in Part (b).

$X_{2}$ = 8.750: The change in mean Rating for Sweetness level 4 relative to sweetness level 2 while keeping Moisture constant is 55.15.

$X_{3}$ = 4.4250: The change in mean Rating for a 1 unit increase in Moisture while keeping Sweetness constant is 4.4250

## 2d. Is the interaction coefficient significant at ð›¼ = 5%? State the null, alternative, decision rule, conclusion.

## \*\*\*2e. your answer is NO in Part (d), drop the interaction term and rerun the model. Write down

the new estimated regression equation at each sweetness level.

# 3. Refer to the "Assessed Valuations" dataset (Value.txt)

```{r}
value_data = read.table(file = "C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Week 10/Week 10/Value.txt", sep= "")

View(value_data)

# #Adding headers
names(value_data) <- c("Price", "Valuation", "Corner_lot")

# names(bank_data) <- c("", "")

#Defining dependent and independent vars
Price = value_data$Price #Y
Valuation = value_data$Valuation #X1
Corner_lot = value_data$Corner_lot #X2
```

## 3a. Regress selling price on lot location only. Write down the estimated regression equation.

```{r}
#Coding Corner Location as a dummy variable
cat_corner_lot <- as.factor(Corner_lot)

#For good measure I set 0 as base reference group
cat_corner_lot <- relevel(cat_corner_lot, ref = "0")

location_onlylm <- lm(Price ~ Corner_lot, data = value_data)
summary(location_onlylm)
```

## 3b. Based on your regression result in Part (a), what is the estimated mean selling price for corner lots? For non-corner lots?

R uses alpha-numeric ordering as the reference group by default so either 'a' or '0' as the base reference group.

Estimated mean selling price for non-corner lots:

**The estimated mean selling price for non-corner lots is roughly \$81K**

For corner lots:

**The estimated mean selling price for corner lots is roughly \$72.5k, which is taken by 81k - 8.5k.**
