---
title: 'STAT 511: HW #7'
author: "Rumil Legaspi"
date: "4/19/2021"
output: pdf_document
---

# Multiple Regression & Brand Preference Dataset

### Setting up workspace

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
library(nortest)
library(olsrr)
library(car)
library(lmtest)
library(MASS)
library(tidyverse)

setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Week 10/Week 10")

brand_data = read.table(file = "Brand.txt", header = FALSE, sep = "")

View(brand_data)

# #Adding headers
names(brand_data) <- c("Rating", "Moisture", "Sweetness")

# names(bank_data) <- c("", "")

#Defining dependent and independent vars
Rating = brand_data$Rating #Y
Moisture = brand_data$Moisture #X1
Sweetness = brand_data$Sweetness #X2
```

# 1a. Regress degree of brand liking on sweetness only. Write down the estimated regression model.

```{r}
#Coding Sweetness as a dummy variable
cat_sweetness <- as.factor(Sweetness)

#Regressing Rating on new Sweetness dummy variable
Sweetness_only <- lm(Rating ~ cat_sweetness, data = brand_data)

summary(Sweetness_only)
```

Estimated Regression model:

$\hat{Y}= 77.375 + 8.750X$

```{r}
#Checking how many levels there are in sweetness as a dummy
levels(cat_sweetness)
typeof(cat_sweetness)
attributes(cat_sweetness)
```

So in this case since sweetness is a category with only 2 level (sweetness level 2 and sweetness level 4) we can think of our regression model like so:

We will use sweetness level 2 as our reference group

```{r}
#the "2" indicates the level marked as "2" given from our level() function and now setting sweetness level 2 as our reference group
cat_sweetness_new <- relevel(cat_sweetness, ref = "2")

#Regressing
Sweetness_only2 <- lm(Rating ~ cat_sweetness_new, data = brand_data)

summary(Sweetness_only2)
```

After doing so we can now say

| $\hat{Y}= B_{0} + B_{1}X_{1}$ | Sweetness = Level 2          | Sweetness = Level 4          |
|-------------------------------|------------------------------|------------------------------|
| $\hat{Y}= 77.375 + 8.750X$    | $\hat{Y}= 77.375 + 8.750(0)$ | $\hat{Y}= 77.375 + 8.750(1)$ |
|                               | = 77.375                     | = 77.375 + 8.750             |
|                               | 77.375                       | 86.125                       |

: Mean Rating at Level 2 or Level 4 Sweetness

```{r}
warnings()
```
