---
title: 'STAT 511: HW #4'
author: "Rumil Legaspi"
date: "1 March 2021"
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 2 
    latex_engine: xelatex
---

## __Workspace Setup__

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
library(nortest)
library(olsrr)
library(car)
library(lmtest)
library(MASS)
library(ggplot2)

setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Week 6/Hw 4")

gpa_data = read.table(file = "GPA.txt", header = FALSE, sep = "")

gpa_data_extended = read.table(file = "GPA_Extended.txt", header = FALSE, sep = "")

# #Adding headers
names(gpa_data) <- c("GPA", "ACT")
names(gpa_data_extended) <- c("GPA", "ACT", "IQ", "Rank")
# names(bank_data) <- c("", "")

#Defining dependent and independent vars
GPA = gpa_data$GPA #Y
ACT = gpa_data$ACT #X1
IQ = gpa_data_extended$IQ #X2
RANK = gpa_data_extended$Rank #X3

gpa_lm = lm(GPA ~ ACT, data = gpa_data)
summary(gpa_lm)
```

 _Refer to the GPA problem (GPA.txt) for Questions 1-3_

# __1. Diagnostic plots:__

## __(a). Plot the regression residuals against the predicted values of the ð‘Œ variable (residuals on the vertical axis). Check the linearity assumption visually.__

```{r}
plot(gpa_lm$fitted.values, gpa_lm$residuals)
```

THe assumption of linearity is __not violated__ because we are __not seeing__ any systemic patterns in the plots.


## __(b). Draw the boxplot, histogram, and normal probability plot of the regression residuals. Check the normality assumption visually.__

```{r}
#boxplot
boxplot(gpa_lm$residuals)

#histogram
hist(gpa_lm$residuals)

#Plotting specifically for QQ Plot
plot(gpa_lm, c(2))
```



Based on the outputs, the box plot is asymmetrical as shown by the outliers, the histogram is left skewed similar to what the normal Q-Q plot is indicating. These plots show that the assumption of normality is violated. 

## __(c). Plot the regression residuals against the ð‘‹ variable (residuals on the vertical axis). Check the equal variance assumption visually.__


```{r}
plot(gpa_data$ACT, gpa_lm$residuals)
```
Based on the plot, we can see no systematic pattern and can therefore conclude visually that our assumption of equal variance
is not violated.



# __2. Diagnostic Tests:__

## __(a). Use normality tests to check the normality assumption and draw a conclusion.__

### Stating our Hypothesis

__Null Hypothesis__: ${H_0}$: The data __is__ from a normal distribution

__Alternative Hypothesis__: ${H_1}$: The data is __NOT__ from a normal distribution
      

### Testing our Hypothesis

__To test these we can use several normality tests using...__

- Shapiro-Wilk normality test
- Shapiro-Francia normality test
- Anderson-Darling normality test

```{r}
#Shapiro-Wilk normality test
shapiro.test(gpa_lm$residuals)

#Shapiro-Francia normality test
nortest::sf.test(gpa_lm$residuals)

#Anderson-Darling normality test
nortest::ad.test(gpa_lm$residuals)
```

## Interpretation of Normality Tests

Looking at the results of these three tests we can see that the p-values are smaller than our alpha. Therefore we reject our NULL hypothesis and that __there is an issue and a violation of our normality assumption.__


## __(b). Use Modified Levene Test and Breusch-Pagan Test to check the equal variance assumption and draw a conclusion.__

## Testing Equal Variance Assumptions

### Stating our Hypothesis

__Null Hypothesis__: ${H_0}$:  The variances in the data __is__ equal
__Alternative Hypothesis__: ${H_1}$: The variances in the data are __NOT__ equal

### Testing our Hypothesis Using Breusch-Pagan test


```{r}
#Conducting Levene Test splitting into two groups

#obtaining median of X to use as a threshold
gpa_median = median(gpa_data$ACT)
#ifelse: spliting X into 2 groups
  #one group x < gpa_median, another with x >= gpa_median
  #ifelse( "if this equation is true", "then do this", "else do this")
gpa_group = ifelse(gpa_data$ACT < gpa_median,
                   "Group1",
                   "Group2")

#Levene "Modified" test using median (default in R)
leveneTest(gpa_lm$residuals, gpa_group)

#bf.test(infection_lm, data = SENIC)

lmtest::bptest(gpa_lm, studentize = FALSE)

```



Both Modified Levene and Breusch-Pagan test gives us high p values indicating that we cannot __reject the null hypothesis__ and conclude that there is no issue with our equal variance assumption.


## __(c). Conduct a lack-of-fit test for the regression model and conclude on the model fitness.__


### Stating our Hypothesis

__Null Hypothesis__: ${H_0}$:  The regression line __IS adequate__ in describing the relationship between ACT and GPA

__Alternative Hypothesis__: ${H_1}$: The regression line is __NOT adequate__ in describing the relationship between ACT and GPA

### Testing our Hypothesis by Conducting our Lack of Fit Test (F Test)

```{r}
#Lack of Fit Test
ols_pure_error_anova(gpa_lm)
```

Based on the small p value 0.003243287, we reject the null hypothesis and conclude with the alternative hypothesis that our regression line is __NOT adequate__ in describing the relationship between ACT and GPA.



# __3. Remediation:__

## __(a). Use Box-Cox method to find the best transformation of ð‘Œ based on a range of ðœ† âˆˆ [âˆ’3, 3], i.e., what is an approximate value of the optimalðœ†in ð‘Œðœ†?__

```{r}
#Applying Box-Cox Method
MASS::boxcox(gpa_lm, lambda = seq(-3, 3, by = 0.1), plotit = FALSE)
MASS::boxcox(gpa_lm, lambda = seq(-3, 3, by = 0.1), plotit = TRUE)
```
 
The graph shows us the most optimal value of $\lambda$ is at roughly 2.1 

## __(b). Plot a smooth curve that best fits the dataset using LOESS method. Is the fitted smooth curve close to linear?__

```{r}
#LOESS scatterplot and smoothed curve
smoothplot = qplot(gpa_data$ACT, gpa_data$GPA, geom=c("point", "smooth"))
smoothplot
```

***The fitted smooth curve is somewhat linear although the fit it is not a good fit.

# __4. Check for Omitted Predictors__










