---
title: 'STAT 511: Assignment #6'
author: "Rumil Legaspi"
date: "4/10/2021"
output: pdf_document
---

# Multiple Regression & Brand Preference Dataset

### Setting up workspace

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
library(nortest)
library(olsrr)
library(car)
library(lmtest)
library(MASS)
library(tidyverse)

setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/Week 10/Week 10")

brand_data = read.table(file = "Brand.txt", header = FALSE, sep = "")

View(brand_data)

# #Adding headers
names(brand_data) <- c("Rating", "Moisture", "Sweetness")

# names(bank_data) <- c("", "")

#Defining dependent and independent vars
Rating = brand_data$Rating #Y
Moisture = brand_data$Moisture #X1
Sweetness = brand_data$Sweetness #X2
```

```{r}
#Regressing Rating (response) on Moisture (explanatory) and Sweetness (explanatory).
#Then summarizing our model
brand_lm <- lm(Rating ~ Moisture + Sweetness, data = brand_data)
summary(brand_lm)
```

## a. Fit a standardized multiple regression model where all variables are centered and scaled.

```{r}
#Scaling coefficients
scaled_Rating <- scale(Rating)
scaled_Moisture <- scale(Moisture)
scaled_Sweetness<-scale(Sweetness)

#putting scaled coefficients into a lm, now the results are scaled, 
scaled_lm <- lm(scaled_Rating ~ 0 + scaled_Moisture + scaled_Sweetness, data = brand_data)

summary(scaled_lm)
```

The estimated intercept coefficient will be zero if all standardized, so we can remove it from our model.

## b. Interpret the partial slope coefficient $\hat{\beta_1}$ in the standardized regression model.

The partial slope coefficient Moisture ($\hat{\beta_1}$) in our standardized regression model can be interpreted like so:

While holding other variables constant and unchanged, when the Moisture variable increases by 1 standard deviation we can expect our response variable, Ratings to increase by 0.89239 standard deviations.

## c. Find the correlation matrix of this dataset. Is there any multicollinearity issue?

```{r}

```

## d. Use the anova() function in R to test if sweetness ($X_{2}$) should be removed from the multiple linear regression, i.e., test the difference between the full model and the reduced model.

```{r}

```
