---
title: 'STAT 511: HW #3 Q:3 & 4'
author: "Rumil Legaspi"
date: "22 February 2021"
output:
  pdf_document:
    toc: yes
    toc_depth: 2 
    latex_engine: xelatex
  html_notebook: default
  html_document:
    df_print: paged
  word_document: default
---

## __Workspace Setup__

```{r, collapse = TRUE, warning = FALSE, message = FALSE}
setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/week 4/Week 4")
#loading gpa data
gpa_data = read.table(file = "GPA.txt", header = FALSE, sep = "")

#Adding headers
names(gpa_data) <- c("GPA", "ACT")

#Defining dependent and independent vars
ACT = gpa_data$ACT #X
GPA = gpa_data$GPA #Y

gpa_lm = lm(GPA ~ ACT, data = gpa_data)

#loading muscle data
muscle_data = read.table(file = "Muscle.txt", header = FALSE, sep = "")

#Adding headers
names(muscle_data) <- c("Muscle", "Age")

#Defining dependent and independent vars
Age = muscle_data$Age #X
Muscle = muscle_data$Muscle #Y

#creating our linear model
muscle_lm = lm(Muscle ~ Age, data = muscle_data)
```
# __3. Refer to the GPA problem in HW #1__


## __(a). Compute the Pearson correlation coefficient and attach the appropriate sign.__


```{r}
#Computing the sample correlation of two variables
#Choose any continuous variable, order does not matter
cor.test(ACT, GPA, method = "pearson")
```
The result of our Pearson correlation coefficient is a positive 0.269. 

__$r = +0.27$__

## __(b). Obtain Spearman rank correlation coefficient.__

```{r, warning = FALSE, message = FALSE}
#spearman correlation
cor.test(ACT, GPA, method = "spearman")
```
The result of our Spearman correlation coefficient is 0.31.


__$r_{s}=0.31$__


## __(c). Which correlation coefficient is stronger? Why?__

The Spearman correlation coefficient is stronger compared to the Pearson correlation between the variables ACT(X) and GPA(Y). This is because Spearman measures the data __points by ranks__ and in this case the relationship in terms of ranking makes for a stronger correlation.


----------------------------------------------

# __4. Refer to the Muscle Mass problem in HW#1.__

## __(a). Compute the Pearson product-moment correlation coefficient.__

```{r}
#Age is x, Muscle is Y. But still placement order does not matter
cor.test(Age, Muscle, method = "pearson")
```
The result of our Pearson correlation coefficient is a negative 0.86 indicating is a strong negative linear relationship between our variables, age and muscle. 

__$r = -0.86$__


## __(b). Based on Part (a), test whether muscle mass and age are significantly correlated in the population at alpha = 0.05. State the null and alternative hypotheses, decision rule, and conclusion.__

- 1. __Pearson Correlation Hypothesis Testing Stating Our Hypotheses:__ 
  + Null hypothesis: $H_{0}$: $\rho = 0$ (There is no statistically significant linear correlation between Age and Muscle)
  + Alternative hypothesis: $H_{1}$: $\rho \neq 0$ (There exists a statistically significant linear correlation between Age and Muscle)


 
```{r}
summary(muscle_lm)

#Helps us find our T value
cor.test(Age, Muscle, method = "pearson")

#Gives us our critical value so we can compare with T statistic
qt(0.975, 58)
```
- 2. Calculating t statistic:
  + $t= -13.193$ 
  + Critical value = 2.00

- 3. Comparing with critical value and p value
  + |$t= -13.193$| > 2.00
  + p value  2.2e-16 < $\alpha$ (0.05)
- 4. Draw conclusion
  + Also, since the p value is very small and we can reject the NULL hypothesis that there is no linear correlation between age and muscle.
  + Therefore we can conclude with the alternative hypothesis that age and muscle are linearly correlated.

## __(c). Compute the Spearman rank correlation coefficient.__
```{r}
#Age is x, Muscle is Y. But still placement order does not matter
cor.test(Age, Muscle, method = "spearman")
```
## __(d). Repeat the test in Part (b) using the Spearman rank correlation from Part (c).__

- 1. __Spearman Correlation Hypothesis Testing Stating our Hypotheses:__ 
  + Null hypothesis: $H_{0}$: $\rho = 0$ (There is no statistically significant rank correlation between Age and Muscle)
  + Alternative hypothesis: $H_{1}$: $\rho \neq 0$ (There exists a statistically significant rank correlation between Age and Muscle)

```{r, warning = FALSE, message = FALSE}
#Helps us find our T value
cor.test(Age, Muscle, method = "spearman")

#Gives us our critical value so we can compare with T statistic
qt(0.975, 58)
```
 
- 2. Calculating t statistic:
  + $t= -13.193$ 
  + Critical value = 2.00

- 3. Comparing with critical value and p value
  + |$t= -13.193$| > 2.00
  + p value  2.2e-16 < $\alpha$ (0.05)
- 4. Draw conclusion
  + Similarly, since the p value is very small we can reject the NULL hypothesis that there is no significant rank correlation between age and muscle.
  + Therefore we can conclude with the alternative hypothesis that there is a significant rank correlation between age and muscle.


## __(e). How do your correlation coefficient estimates and test conclusions in Parts (a) and (b) compare to those obtained in Parts (c) and (d)?__

Based on the results from our hypothesis tests on both Pearson and Spearman, both age and muscle hold a significant rank and negative linear correlation relationship. These two hypothesis help verify our correlation coefficient estimates and allow us to safely conclude that a correlation does exist.












