---
title: 'STAT 511: HW #3 Q:1 & 2'
author: "Rumil Legaspi"
date: "15 February 2021"
output:
  pdf_document: 
    latex_engine: xelatex
  html_notebook: default
  html_document:
    df_print: paged
  word_document: default
---





```{r}
setwd("C:/Users/RUMIL/Desktop/APU/STAT 511 - Millie Mao (Applied Regression Analysis)/week 4/Week 4")


gpa_data = read.table(file = "GPA.txt", header = FALSE, sep = "")

#Adding headers
names(gpa_data) <- c("GPA", "ACT")
head(gpa_data)

#Defining dependent and independent vars
ACT = gpa_data$ACT #X
GPA = gpa_data$GPA #Y

gpa_lm = lm(GPA ~ ACT, data = gpa_data)
summary(gpa_lm)

```

size.lm = lm(hours~size, data = lotdata)
summary(size.lm)

#ANOVA decomp[osition
aov(size.lm)

#ANOVA with F-test
anova(size.lm)

#General linear test
red.lm = lm(hours~1, data = lotdata) #hours~1 is reduced model
anova(red.lm, size.lm)   

## __1.Refer to the GPA problem in HW#1__


### __(a).Setting up the ANOVA table__ 

```{r}
anova(gpa_lm)
```


### __(b). What does MSR measure in your ANOVA table? What does MSE measure? Under what condition do MSR and MSE estimate the same quantity?__

To begin with MSR (Mean Squared of Regression) comes from the sum of squares (***of our X values) divided by the degrees of freedom from our regression model. In this case MSR measures our regression model's variability (separate from the variability of our error(MSE)).

MSE is the mean squared error and measures how close our regression line is to the data points. It measures the distances from the data points to our regression line. Those distances are the errors and the MSE is squaring those errors. __Ultimately__ it measures the variability/spread of those errors.


Both the MSR and MSE estimate the same quantity when the slope  $\beta_{1}$ is zero or not.
(*** from the slide 8/39, the same quantity error is the same quantity measured when slope is zero or not)

### __(c). at $\alpha$ = 0.05, conduct an F-test of whether or not $\beta_{1}$ = 0. State the null and alternative hypotheses, decision rule, and conclusion.__

<div align="center">

Null hypothesis: H0: $\beta_{1}$ = 0 (slope is horizontal/ no relationship)

Alternative hypothesis: H1: $\beta_{1}$ /= 0 (slope exists/ relationship exists)
      
</div>

Given our F-value is 9.2402 

__note: 1. If the f statistic is larger than the critical value 

2. or the p value is less than alpha then it is significant and we reject the null__
```{r}
#Using qt() to find our critical value we get
qt(0.995, 118)
#*** why do we use 0.995 and not 0.99 here?
```

The F-value is greater than our critical value (2.618137) __AND__ our p-value (0.002917) is less than our $\alpha$ = 0.05 we reject our null and therefore a non-zero slope exists and a significant relationship between ACT(X) and GPA(Y) scores.




##currently at 1:07:00 in class recording


















